\section{EXI Options}
\label{sec:Options}

The EXI specification defines a set of options that can be used to alter the way the
EXI body is processed. The precise values of the EXI options must be known by the
EXI processors to correctly interpret the EXI streams. One way to communicate the
options used for processing is to include them in the EXI Header (Recommended).
However, the presence of the EXI Options in its entirety in the EXI header is
optional according to the EXI specification. This means that if the options are not
included in the EXI header they must be communicated with an out-of-band mechanism.

During encoding, the options are included in the EXI Header if the \texttt{has\_options}
flag is set to TRUE:
\begin{lstlisting}
strm.header.has_options = TRUE;
\end{lstlisting}
If the flag is not set to TRUE, the specified options are used for encoding and
must be known (communicated out-of-band) by the decoder as they are not included
in the header.

During decoding, if the options are not present in the EXI Header they must be
known and set before the parsing of the header with \texttt{parseHeader()}. If
out-of-band options are set for decoding and there are options specified in the
EXI stream - the out-of-band options will be ignored. If there is no value
specified for a particular option in the EXI Header or by out-of-band
mechanism then the default value for that option is assumed
according to the EXI specifications.

Table \ref{tab:Options} shows how to set the Header fields and each of the EXI
options before processing EXI stream with EXIP.

\begin{table}[ht]\footnotesize
\centering
\begin{tabular}{l p{9.5 cm}}
\toprule
\textbf{EXI Option / Header field} & \textbf{Instructions} \\
\midrule
\midrule
EXI Cookie                &  \texttt{strm.header.has\_cookie = TRUE;} (Default is FALSE)\\
Options Presence Bit      &  \texttt{strm.header.has\_options = TRUE;} (Default is FALSE)\\
EXI Version               &  \texttt{strm.header.version\_number = 2;} (Default is 1)\\
\midrule
alignment                 &  \texttt{SET\_ALIGNMENT(strm.header.opts.enumOpt, V);} (Where V in \{\texttt{BIT\_PACKED, BYTE\_ALIGNMENT, PRE\_COMPRESSION}\})\\
compression               &  \texttt{SET\_COMPRESSION(strm.header.opts.enumOpt);}\\
strict                    &  \texttt{SET\_STRICT(strm.header.opts.enumOpt);}\\
fragment                  &  \texttt{SET\_FRAGMENT(strm.header.opts.enumOpt);}\\
preserve                  &  \texttt{SET\_PRESERVED(strm.header.opts.preserve, V);} (Where V in \{\texttt{PRESERVE\_COMMENTS, PRESERVE\_PIS, PRESERVE\_DTD, PRESERVE\_PREFIXES, PRESERVE\_LEXVALUES}\})\\
selfContained             &  \texttt{SET\_SELF\_CONTAINED(strm.header.opts.enumOpt);}\\
schemaId                  &  \texttt{const String SCHEMA\_ID\_STR = \{"http://www.ltu.se/exip", 22\};}
\texttt{strm.header.opts.schemaID = SCHEMA\_ID\_STR;}\\
datatypeRepresentationMap &  Not implemented yet!\\
blockSize                 &  \texttt{strm.header.opts.blockSize = 1000;}\\
valueMaxLength            &  \texttt{strm.header.opts.valueMaxLength = 100;}\\
valuePartitionCapacity    &  \texttt{strm.header.opts.valuePartitionCapacity = 100;}\\
{[user defined meta-data]}  &  Not implemented yet!\\
\bottomrule
\end{tabular}
\caption{Instructions on how to set each EXI options before processing}
\label{tab:Options}
\end{table}