\section{Serialization}
\label{sec:Serialization}

This section describes the process of serializing an EXI stream using the EXIP API.
The description is divided into two parts - schema-less and schema-enabled encoding.
The basic steps are essentially the same in both cases so the schema-enabled part
shows only some special use-cases and configurations that differ from the schema-less
serialization. 

As discussed in section \ref{sec:Basic-Concepts}, the EXIP API is especially useful
for applications that are not dependent on XML APIs and text XML inputs. Example
use cases are embedded devices working exclusively with EXI, XML binding tools
that generate parsing/serialization code from XML Schema definitions, productivity tools
etc. If your application works with text XML input that needs to be converted to EXI
you need to create an external module that is built on top of the EXIP serialization API.

The serialization API is defined in \texttt{EXISerializer.h}. You can work with the
serialization functions directly or use the \texttt{serialize} global variable to
access these functions as methods.
 
\subsection{Schema-less encoding}
The serialization of an EXI stream consists of 7 simple steps:
\begin{enumerate}
 \item Declare a stream container that holds the serialized data and EXIP state:
\begin{lstlisting}
EXIStream strm;
\end{lstlisting}

 \item Initialize the EXI header of the new stream container:
\begin{lstlisting}
serialize.initHeader(&strm);                                                            
\end{lstlisting}

 \item \label{item:step-options} (Optional) Set any options in the EXI header, if different from the defaults:
\begin{lstlisting}
strm.header.has_options = TRUE; /* Set the options presence bit */ 
strm.header.has_cookie = TRUE; /* Include EXI cookie in the header */
SET_STRICT(strm.header.opts.enumOpt); /* Indicate a STRICT mode encoding */
\end{lstlisting}
(see Section \ref{sec:Options} for more information on how to set up
the EXI options)

 \item Define a binary output buffer for storing the serialized EXI data.
The buffer can optionally include an external stream for sending the
serialized output (flushing the buffer when full). If an output stream
is not defined the whole EXI document must fit into the memory buffer.
The following code declares and initializes an output memory buffer:
\begin{lstlisting}
BinaryBuffer buffer;
char buf[OUTPUT_BUFFER_SIZE];

buffer.buf = buf;
buffer.bufLen = OUTPUT_BUFFER_SIZE;
buffer.bufContent = 0;
\end{lstlisting}
Beside these definitions, when a file is used to store the
serialized data the following steps are also required: 
\begin{lstlisting}
/* IN CASE OF FILE OUTPUT STREAM */
/* Define a function implementing the actual writing to a file */
size_t writeToFileOutputStream(void* buf, size_t writeSize, void* stream)
{
	FILE *outfile = (FILE*) stream;
	return fwrite(buf, 1, writeSize, outfile);
}

FILE *outfile; /* Using a file for storing the serialized data */
outfile = fopen(destinationEXIFile, "wb" ); /* open the file before use */
buffer.ioStrm.readWriteToStream = writeToFileOutputStream;
buffer.ioStrm.stream = outfile; /* Sets the output stream to the file */
\end{lstlisting}
If no file or other output streams are being used (e.i. the whole EXI message is
stored in \texttt{buffer.buf}), the output function and the stream
of the \texttt{BinaryBuffer} must be initialized with NULL values:
\begin{lstlisting}
/* IN CASE OF IN-MEMORY ONLY OUTPUT*/
buffer.ioStrm.readWriteToStream = NULL;
buffer.ioStrm.stream = NULL;
\end{lstlisting}

The size of the \texttt{OUTPUT\_BUFFER\_SIZE} macro depends on the
use case. When in in-memory output mode or sufficient RAM on the
platform the buffer size should be kept high. Note that even if
an output stream is used such as a file, a non-zero length buffer
for temporary storing parts of the EXI serialization is still needed.

 \item \label{item:initialization} Initialize the stream:
\begin{lstlisting}
/**
 * @param[in, out] strm EXI stream container
 * @param[in, out] buffer output BinaryBuffer for storing the encodded EXI stream
 * @param[in] NULL a compiled schema information to be used for schema
		      enabled processing; NULL if no schema is available
 * @param[in] SCHEMA_ID_ABSENT one of SCHEMA_ID_ABSENT, SCHEMA_ID_SET,
		SCHEMA_ID_NIL or SCHEMA_ID_EMPTY as defined in the specification
 * @param[in] NULL when in SCHEMA_ID_SET mode a valid string representing
		      the schemaID; NULL otherwise
 */
serialize.initStream(&strm, buffer, NULL, SCHEMA_ID_ABSENT, NULL);                      
\end{lstlisting}

 \item \label{item:encoding} Start building the stream step by step starting from the header. The body should always start with
    \texttt{startDocument()} and ends with \texttt{endDocument()} with at least one element declaration.
In schema-less mode the only type of data allowed is \texttt{string} so when encoding element or
attribute values you should use \texttt{stringData()}. \linebreak 
\textbf{NOTE:} different attributes within a single element should be encoded in a lexicographical order during serialization!
\begin{lstlisting}
/* Static String type definitions if any */
const String NS_TARGET_STR = {"http://www.ltu.se/exip", 22};
const String NS_EMPTY_STR = {NULL, 0};
const String ELEM_ROOT_STR = {"rootElement", 11};
const String ATTR_TEST_STR = {"testAttribute", 13};

serialize.exiHeader(&strm); /* Always first */
serialize.startDocument(&strm); /* Indicates beginning of the EXI body */

QName qname= {&uri, &ln, NULL}; /* QName definition cosisting of
namespace, local name and prefix if any*/

/* Encode an element with QName <http://www.ltu.se/exip:rootElement> */
qname.uri = &NS_TARGET_STR;
qname.localName = &ELEM_ROOT_STR;
serialize.startElement(&strm, qname);

/* Encode attribute with QName <testAttribute> and indicate its type */
qname.uri = &NS_EMPTY_STR;
qname.localName = &ATTR_TEST_STR;
serialize.attribute(&strm, qname, VALUE_TYPE_STRING);

String ch; /* EXIP string representing a string value */

/* Convert a static ASCII string constant to EXIP string type */
asciiToString("attribute value", &ch, &strm.memList, FALSE);

serialize.stringData(&strm, ch); /* Encode the value of the attribute */

asciiToString("element value", &ch, &strm.memList, FALSE);

serialize.stringData(&strm, ch); /* The value of the <rootElement> */

serialize.endElement(&strm); /* Close element <rootElement> */

serialize.endDocument(&strm); /* Close the EXI body*/
\end{lstlisting}

The above code produces an EXI stream that has the following XML representation:
\begin{lstlisting}
<?xml version="1.0" encoding="UTF-8"?>
<xp:rootElement xmlns:xp="http://www.ltu.se/exip"
		testAttribute="attribute value">
element value
</xp:rootElement>
\end{lstlisting}

 \item Destroy the EXI stream container and free the memory allocated by it. If any other
streams are left open close them as well:
\begin{lstlisting}
serialize.closeEXIStream(&strm);
fclose(outfile); /* Close the file (if any) used to store the output data */
\end{lstlisting}

\end{enumerate}


\subsection{Schema-enabled encoding}
When in schema-enabled mode the basic serialization steps are essentially the same.
The difference is in the parameters passed to the \texttt{serialize.initStream()}.
You need to pass a valid \texttt{EXIPSchema} object and not NULL as in the schema-less
case. This object contains all the definitions and constrains from the XML schema and
its creation is a topic of section \ref{sec:Schema-Infromation} \textbf{Schema Information}.
During the encoding in Step \ref{item:encoding} you must use the types for the
values as defined by the schema definitions. For example, \texttt{serialize.intData()} for
integer values of elements or attributes and \texttt{serialize.booleanData()} for boolean values.
There are also few options
in the EXI header that influence how the schema information is utilized when encoding
the EXI body:
\begin{description}
 \item[strict] this is a boolean option that specify if deviations from the schema
definitions are allowed (has value of \texttt{false}) or are not allowed (has value of \texttt{true})
in the EXI body. The default value of this option is \texttt{false}. When in STRICT mode
the representation is slightly more compact. You can set the
value of this option in Step \ref{item:step-options} with the following macro:
\begin{lstlisting}
SET_STRICT(strm.header.opts.enumOpt); /* Indicate a STRICT mode encoding */
\end{lstlisting}

 \item[schemaId] this option takes a string for its value. There are four possible states
for this option as defined in the \href{http://www.w3.org/TR/2011/REC-exi-20110310/#key-schemaIdOption}{specification}:
\begin{itemize}
 \item SCHEMA\_ID\_ABSENT - default state; no statement is made about the schema information.
If it is a schema-less or schema-enabled mode and what schema definitions to be used when decoding must be communicated out of band.
 \item SCHEMA\_ID\_SET - a string identification of the schema used for encoding is specified in the EXI header.
The format of this schema identifier is not standardized and is left for the application to decide what to use.
One possibility is to use the target namespace of the schema or the URL address of the schema location.
 \item SCHEMA\_ID\_NIL - no schema information is used for processing the EXI body (i.e. a schema-less EXI stream).
 \item SCHEMA\_ID\_EMPTY - no user defined schema information is used for processing the EXI body; however,
the built-in XML schema types are available for use in the EXI body
through the \texttt{xsi:type} attribute.
\end{itemize}
  All of these states can be set during initialization - Step \ref{item:initialization}:
\begin{lstlisting}
String schemaID;
asciiToString("http://www.ltu.se/schema", &schemaID, &strm.memList, FALSE);
serialize.initStream(&strm, buffer, OUTPUT_BUFFER_SIZE, &out, NULL, 
					SCHEMA_ID_SET, schemaID);                           
\end{lstlisting}
In this example we defined the state to SCHEMA\_ID\_SET and set an identifier for the schema used - \texttt{http://www.ltu.se/schema}.
 \end{description}

The serialization interface defined in \texttt{EXISerializer.h} has a more efficient low level API
for encoding the document structure when in schema-enabled mode. This API can be accessed through \texttt{serializeEvent()}
function. A description on how it can be used might be included in the future versions of this guide. If you want to know more about it
you can check how it is used to encode the EXI options document in the header of an EXI stream in \texttt{headerEncode.c}. 